---
title: "Budapest Airport Traffic Data"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r options, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message=FALSE)
options(xtable.comment = FALSE)
```

```{r setup, include=FALSE, message=FALSE}
if(!require(flexdashboard)){install.packages("flexdashboard")}
if(!require(leaflet)){install.packages("leaflet")}
if(!require(leaflet.extras)){install.packages("leaflet.extras")}
if(!require(htmltools)){install.packages("htmltools")}
if(!require(readr)){install.packages("readr")}
if(!require(tidyverse)){install.packages("tidyverse")}
if(!require(plotly)){install.packages("plotly")}
if(!require(DT)){install.packages("DT")}
if(!require(ggplot2)){install.packages("ggplot2")}
if(!require(xts)){install.packages("xts")}
if(!require(dygraphs)){install.packages("dygraphs")}
if(!require(ggthemes)){install.packages("ggthemes")}
if(!require(xtable)){install.packages("xtable")}
if(!require(knitr)){install.packages("knitr")}
if(!require(crosstalk)){install.packages("crosstalk")}
```

```{r download data}
dataviz <- read.csv("data/2013_2017_monthly_BUD_traffic.csv", header=TRUE, sep = ";")

#unic destinations with overall summary data and with lon and lat coordinates
destinations <-read.delim("data/destinations.txt", header=TRUE, sep = ",")

#dataviz with lon and lat coordinates
dataviz.lon.lat <-read.delim("data/dataviz_lon_lat.txt", header=TRUE, sep = ",")
```

```{r dataviz.destinations}

dataviz.destinations <- dataviz.lon.lat  %>%
  filter(Country != "Unknown" ) %>%
  group_by(City, Country, lon,lat) %>%
  summarise( ATM = sum(ATM),
             PAX  = sum(PAX),
             Capacity = sum(Capacity),
             potential = round(PAX/Capacity * 100),
             seats = round(mean(PAX/ATM))) %>%
  arrange(desc(PAX))

dataviz.destinations <- cbind(rank = as.numeric(rownames(dataviz.destinations)), dataviz.destinations) # add rowid as first column

```


```{r top destinations by year}

dataviz.destinations.year <- dataviz.lon.lat  %>%
  filter(Country != "Unknown" ) %>%
  group_by(Year, City, Country, lon, lat) %>%
  summarise( ATM = sum(ATM),
             PAX  = sum(PAX),
             Capacity = sum(Capacity),
             potential = round(PAX/Capacity * 100),
             seats = round(mean(PAX/ATM))
            ) %>%
  arrange(desc(PAX,Year))

#add a summary rank by year and PAX
dataviz.destinations.year <- dataviz.destinations.year %>% arrange(desc(PAX,Year)) %>%
    group_by(Year)%>%
mutate(rank = rank(-PAX, ties.method = "first"))%>%
ungroup


```


```{r top destinations per year in table format}
# TO DO :: in more general format 
top.dest.2017 <- dataviz  %>%
  filter(Country != "Unknown", Year =='2017' ) %>%
  group_by(Year, City, Country) %>%
  summarise( ATM = sum(ATM),
             PAX  = sum(PAX),
             Capacity = sum(Capacity),
             potential = round(PAX/Capacity * 100),
             seats = round(mean(PAX/ATM))) %>%
  arrange(desc(PAX,Year))

# create am empty df
df_result <-data.frame(City = character(), dif_PAX = numeric(), dif_potential = numeric(), dif_seats=numeric())
df_result <- df_result[-1,]

# search the relevant values from prev
for (city in top.dest.2017[1:10,]$City){
my_df_2017 <- dataviz.destinations.year[ which(dataviz.destinations.year$Year=='2017'& dataviz.destinations.year$City==city), ]
 
# TO DO :: ifelse for newly added cities
my_df_2016 <- dataviz.destinations.year[ which(dataviz.destinations.year$Year=='2016'& dataviz.destinations.year$City==city), ]

# calculating the percentage change from year 2016
  df_result <- rbind(df_result, 
                     data.frame(City = city, 
              dif_PAX = round((my_df_2017$PAX - my_df_2016$PAX)/my_df_2016$PAX*100),
              dif_potential = round((my_df_2017$potential - my_df_2016$potential)/my_df_2016$potential*100),
              dif_seats = round((my_df_2017$seats - my_df_2016$seats)/my_df_2016$seats*100))
          )
  }

top.dest.table <- add_column(top.dest.2017[1:10,], dif_PAX = df_result$dif_PAX, dif_potential = df_result$dif_potential, dif_seats = df_result$dif_seats)


```


```{r data for yearly trends}

dataviz.trend <- dataviz  %>%
  filter(Country != "Unknown" ) %>%
  group_by(Year) %>%
  summarise( 
             PAX  = sum(PAX),
             Capacity = sum(Capacity)) %>%
  arrange(Year)

#gather.dataviz.trend <- dataviz.trend %>% gather(params, values, -Year)
```


```{r data for monthly trends}

dataviz.month <- dataviz  %>%
  filter(Country != "Unknown" ) %>%
  group_by(Month) %>%
  summarise( 
             PAX  = round(sum(PAX)/5),
             Capacity = round(sum(Capacity)/5)) %>%
  arrange(Month)

#gather.dataviz.month <- dataviz.month %>% gather(params, values, -Month)
```

```{r data for ATM, PAX}

dataviz.ATM.PAX <- dataviz  %>%
  filter(Country != "Unknown" ) %>%
  group_by(Year) %>%
  summarise( 
             ATM  = sum(ATM),
             PAX = sum(PAX)) %>%
  arrange(Year)

```

Overview
====================================================

Column {data-width=400}
-----------------------------------------------------------------------

### Top destinations (total passenger traffic)

```{r}
plot.destinations.year <- ggplot(dataviz.destinations.year[1:50,], 
                            aes( x = reorder(City, PAX, min), y = PAX), color=City) +
  geom_bar(stat = "identity", position = "dodge", width = .4)+
  labs(title = "Top 10 destinations", 
       subtitle = "",
       x = "", 
       y = "")+
  theme_economist_white(gray_bg = FALSE) +
    scale_colour_economist()+
  theme(axis.text.x=element_text(angle=30,hjust=1,vjust=0.5, size = 10, face = "bold"),
        axis.text.y=element_text(face = "bold"),
        plot.title=element_text(size = 24)) +
  coord_flip()+
  facet_wrap( ~ Year,nrow = 1)

plot.destinations.year
```

### Year-over-year growth for 2017, compared to 2016

```{r}

top.dest.table <- top.dest.table %>%
  ungroup(top.dest.table) %>%
  select(City, Country, PAX, dif_PAX) %>%
  mutate(dif_PAX = paste(dif_PAX, "%", sep=""))

kable(top.dest.table, knit.table.format = "html", row.names = FALSE, col.names = c("City", "Country", "Passangers", "growth"))

```

```{r}

```


Column {data-width=350}
-----------------------------------------------------------------------

### Year-over-year growth of passangers and capacity for 2013-2017

```{r - trends in PAX}
dataviz.date <- dataviz  %>%
  filter(Country != "Unknown" ) %>%
  mutate(Month = ifelse( Month %in% c(1,2,3,4,5,6,7,8,9), paste("0", Month, "-01",sep=""), paste( Month, "-01",sep=""))) %>%
  unite(Datum,1:2, sep="-")

dataviz.date <- dataviz.date  %>%
  filter(Country != "Unknown" ) %>%
  group_by(Datum) %>%
  summarise( 
             
             PAX = sum(PAX),
             Capacity = sum(Capacity)) %>%
  arrange(Datum)

dataviz.date.ts <- ts(dataviz.date[, -1], frequency=12, start=c(2013,1))

dygraph(dataviz.date.ts, main = "Passangers") %>% 
  dyRangeSelector(dateWindow = c("2013-06-01", "2017-06-01"))
  
```

### Year-over-year aircraft movements trend for 2013-2017

```{r plot for ATM}
dataviz.date <- dataviz  %>%
  filter(Country != "Unknown" ) %>%
  mutate(Month = ifelse( Month %in% c(1,2,3,4,5,6,7,8,9), paste("0", Month, "-01",sep=""), paste( Month, "-01",sep=""))) %>%
  unite(Datum,1:2, sep="-")

dataviz.date <- dataviz.date  %>%
  filter(Country != "Unknown" ) %>%
  group_by(Datum) %>%
  summarise( 
             ATM = sum(ATM)) %>%
  arrange(Datum)

dataviz.date.ts <- ts(dataviz.date[, -1], frequency=12, start=c(2013,1))

dygraph(dataviz.date.ts, main = "Aircrafts") %>% 
  dyRangeSelector(dateWindow = c("2013-06-01", "2017-06-01"))


```

Column {data-width=250, data-height=100}
-----------------------------------------------------------------------


### Total number of passangers 2017

```{r}

trend.2017 <- dataviz.trend[ which(dataviz.trend$Year=='2017'), ]

valueBox(trend.2017$PAX, 
         caption = '',
        #href = "#trip-duration"
         icon = 'fa-pencil'
         )


```



### Top destination from 181 in 2017

```{r}

trend.2017 <- dataviz.trend[ which(dataviz.trend$Year=='2017'), ]

valueBox('London', 
         caption = '',
        #href = "#trip-duration"
         icon = "fa-pencil"
         )


```


### Average number of passangers in monthly basis 2013-2017

```{r - monthly plot}
plot.month <- ggplot(data = dataviz.month) +
  #geom_line(aes(y = PAX, x = Month),  size = 1.2)+
  geom_bar(aes(y = PAX, x = Month), stat = "identity", position = "dodge", width = .4)+
  scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9,10,11,12), labels = c("Jan","Feb","Mar","Apr","May","Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) +
  scale_y_continuous(labels = scales::comma)+
  labs(title ="Monthly trends",
       subtitle = "",
       x = "", 
       y = "")+
  theme_economist_white(gray_bg = FALSE) +
    scale_colour_economist()+
  theme(axis.text.x=element_text(angle=30,hjust=1,vjust=0.5, size = 12, face = "bold"),
        axis.text.y=element_text(face = "bold", size = 12),
        plot.title=element_text(size = 24)) +
  coord_flip()
  
  
plot.month
```

By country
====================================================

Column {data-width=200 .sidebar}
---------------------------------------------------------------------------------

### Countries

```{r}

selectInput("countryId", "Country:",
                sort(unique(dataviz.lon.lat$Country)))

selectInput("yearId","Year:", sort(unique(dataviz.lon.lat$Year)))

country.year.df <- reactive({

  dataviz.lon.lat %>%
    filter(Country == input$countryId, Year==input$yearId)%>%
    mutate(PAX_pct = round(PAX / sum(PAX),4)*100,
           ATM_pct = round(ATM / sum(ATM),4)*100)%>%
    group_by(City,Year,lon,lat) %>%
    summarise(PAX = sum(PAX), 
              Capacity=sum(Capacity),
              ATM=sum(ATM),
              PAX_pct = sum(PAX_pct),
              ATM_pct = sum(ATM_pct))

             
})


```



Column {data-width=450}
-----------------------------------------------------------------------

### Destinations on the map

```{r}
renderLeaflet({
  

df <- country.year.df()

    df %>% leaflet() %>%
      addTiles(group = "OSM") %>% 
      addProviderTiles("CartoDB", group = "CartoDB") %>% 
      addProviderTiles("Esri", group = "Esri") %>% 
      addProviderTiles("Esri.WorldImagery", group = "ESRI World Imagery") %>%
      # Use addLayersControl to allow users to toggle between basemaps
      addLayersControl(baseGroups = c("OSM", "CartoDB", "Esri", "ESRI World Imagery")) %>%
      addCircleMarkers(
          ~ lon ,
          ~ lat,
          popup =  ~(paste("<b>",df$City,"</b>","<br>","Passagers:",df$PAX, "<br>","Flights:", df$ATM)),
          radius = 5,
          stroke = FALSE,
          fillOpacity = 0.8,
          popupOptions = popupOptions(closeButton = FALSE)
  )
})

```



Column {data-width=350}
-----------------------------------------------------------------------

### Cities, passangers %, aircrafts % in the destination country

```{r}

df <- reactive({country.year.df() %>%
  ungroup %>%
  select(City, PAX, PAX_pct, ATM, ATM_pct) %>%
  mutate(PAX_pct = paste(PAX_pct, "%", sep=""),
         ATM_pct = paste(ATM_pct, "%", sep=""))
})

renderTable(df(), colnames = FALSE, striped = TRUE, bordered = TRUE, hover = TRUE)
```


### Number of passangers
```{r}
renderPlot({ 
  ggplot(country.year.df(), aes(x = City,y = PAX)) +
  geom_bar(stat='identity', position = "dodge", width = 0.2) +
    scale_y_continuous(labels = scales::comma) +
    labs( 
      x = "", 
      y = "")+
    theme_economist_white(gray_bg = FALSE) +
    theme(
    axis.text.x=element_text(angle=45,hjust=1,vjust=0.5, size = 8, face = "bold")
    )
  
  }
    )

```


### Aircrafts movements

```{r}
renderPlot({ 
   ggplot(country.year.df(), aes(x = City,y = ATM)) +
  geom_bar(stat='identity', position = "dodge", width = 0.2) +
    labs( 
      x = "", 
      y = "")+
    theme_economist_white(gray_bg = FALSE) +
    theme(
    axis.text.x=element_text(angle=45,hjust=1,vjust=0.5, size = 8, face = "bold")
    )
 
  })
```


Top destinations
====================================================

Column {data-width=300 .sidebar}
-----------------------------------------------------------------------
### Top destinations 

```{r crosstalk}
# using crosstalk

#reorder the columns
dataviz.destinations.year <- dataviz.destinations.year[c('rank','Year','City','Country','PAX','ATM','Capacity', 'potential', 'seats', 'lon','lat')]

# wrap data for crosstalk
sd <-  SharedData$new(dataviz.destinations.year)

filter_select("YearId", "Year", sd, ~Year)

filter_slider("dest_range_sliderId", label = "Select the top destinations range:", sd, column=~rank, step = 1, dragRange = TRUE, width=300)


```

Column {data-width=400}
-----------------------------------------------------------------------

### On the map

```{r}
# Use SharedData like a dataframe with Crosstalk-enabled widgets

  sd %>% 
    leaflet() %>% addTiles() %>% 
          addCircleMarkers(
          ~ lon ,
          ~ lat,
          popup =  ~(paste("<b>",dataviz.destinations.year$City,"</b>","<br>","Passagers:",dataviz.destinations.year$PAX, "<br>","Flights:", dataviz.destinations.year$ATM)),
          radius = 5,
          stroke = FALSE,
          fillOpacity = 0.8,
          popupOptions = popupOptions(closeButton = FALSE)
  )

```

Column {data-width=300}
-----------------------------------------------------------------------

### On table

```{r}
datatable(sd, extensions="Scroller", style="bootstrap", class="compact", width="100%",
    options=list(deferRender=TRUE, scrollY=300, scroller=TRUE), rownames = FALSE)
```


Detailed info
====================================================

### Search

```{r}
datatable(dataviz, rownames = FALSE,
extensions = 'Buttons',
list(
dom = 'Bfrtip',
buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
)
)
```

