---
title: "Preprocess airport data"
output: pdf_document
---

## Codebook
Year, Month: 
Country, City: destination
ATM (AIR Traffic Movement): taking offs and landings
PAX: nr off passangers
Capacity: available nr of seats

## Preprocessing steps
For interactive shiny leaflet it is necessary to add latitude and longitude coordinates to the default data.
1. Major cities coordinates can be added using the `geocode` function from the `ggmap` library.
2. wordcities.txt is downloaded from [MaxMind](https://www.maxmind.com/en/free-world-cities-database)
This step is necessary as no all the cities coordinates are provided by `geocode`.
3. All the destinations are obtained from the initial df.
4. For all the destinations the coordinates are set using the `geocode` or wordcities.txt.
5. Finally, the ccordinates are added to the original dataset.

Files needed for preprocess are in `prepdata`:      
`worldcities.txt`     
`2013_2017_monthly_BUD_traffic.csv` 

The resulted files needed for visualisation are copied in `data`:     
`2013_2017_monthly_BUD_traffic.csv`       
`dataviz_lon_lat.txt`      
`destinations.txt`        



```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message=FALSE)
```

```{r load, message = FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
if(!require(scales)){install.packages("scales")}
if(!require(forcats)){install.packages("forcats")}
#if(!require(data.table)){install.packages("data.table")}
if(!require(knitr)){install.packages("knitr")}
if(!require(tidyr)){install.packages("tidyr")}
if(!require(reshape2)){install.packages("reshape2")}
if(!require(tibble)){install.packages("tibble")}
if(!require(xts)){install.packages("xts")}
if(!require(dygraphs)){install.packages("dygraphs")}
if(!require(ggthemes)){install.packages("ggthemes")}
if(!require(ggmap)){install.packages("ggmap")}

```


```{r - load data }
dataviz <- read.csv("prepdata/2013_2017_monthly_BUD_traffic.csv", header=TRUE, sep = ";")
# check for missing values

sapply(dataviz, anyNA)

```

```{r destinations}
dataviz.destinations <- dataviz  %>%
  filter(Country != "Unknown" ) %>%
  group_by(City, Country) %>%
  summarise( ATM = sum(ATM),
             PAX  = sum(PAX),
             Capacity = sum(Capacity),
             potential = round(PAX/Capacity * 100),
             seats = round(mean(PAX/ATM))) %>%
  arrange(desc(PAX))

write.table(dataviz.destinations, "destinations.txt", row.names = FALSE, sep =",", col.names = TRUE)

```


```{r - worldcities}
# Read tabular data into R
worldcities <-read.delim("prepdata/worldcitiespop.txt", header = TRUE, sep = ",", dec=".")

head(worldcities)
```


```{r}
# get the assumed missing data from worldcities         
worldcities.dest <- worldcities %>%
  filter(AccentCity %in% dataviz.destinations$City)%>%
  group_by(AccentCity, Country)%>%
  mutate(total = n())%>%
  arrange(AccentCity)%>%
  ungroup()%>%
  filter(!is.na(Population))

write.table(worldcities.dest, "worldcities_2017.txt", row.names = FALSE, sep =",", col.names = TRUE)
```

```{r}
#worldcities_2017.txt + prepared by hand
prepare.2017 <-read.delim("worldcities_2017.txt", header=TRUE, sep = ",")
#missing.2017 <-read.delim("missing.txt", header=TRUE, sep = ",")
```

```{r}
# search for the missing list (prepared by hand)
missing.dest <- dataviz.destinations %>%
  filter(!(City  %in% prepare.2017$AccentCity)) %>%
  arrange(City)

write.table(missing.dest, "missing_2017.txt", row.names = FALSE, sep =",", col.names = TRUE)  

```


```{r}

#bind geocode to df
dataviz.destinations <- bind_cols(dataviz.destinations, geocode(as.character(dataviz.destinations$City),  source = c("dsk")))

# for NAs take lat and lon from prepare.2017 df prepared by hand
destinations <- dataviz.destinations %>%
  mutate(lon = ifelse(is.na(lon), prepare.2017$Longitude, lon),
         lat = ifelse(is.na(lat), prepare.2017$Latitude, lat))

# check for missing values
sapply(destinations, anyNA)

write.table(destinations, "data/destinations.txt", row.names = FALSE, sep =",", col.names = TRUE)

# add lon,lat to original df
dataviz.lon.lat <- left_join(dataviz,destinations[c("City","lon","lat")], by ="City") %>%
  filter(Country != "Unknown" )

write.table(dataviz.lon.lat, "data/dataviz_lon_lat.txt", row.names = FALSE, sep =",", col.names = TRUE)


```


```{r}
sapply(dataviz.lon.lat, anyNA)

```


```{r}
#rsconnect::setAccountInfo(name='szintakacseva',
#			  token='1AFB4DF82AC496A2C2D8F858AF759D88',
#			  secret='Fc2pGpkaYZZQuYrGhL09DoRDiKvNhgXWplaTwmWm')
```

